---
title: "Exploratory Analysis"
author: "Ganesh Krishnan"
date: "December 20, 2015"
output: html_document
---

```{r global_opts, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set (echo=FALSE, message=FALSE, warning=FALSE)
```

```{r loadLibraries}
library (plyr)
library (dplyr)
library (ggplot2)
library (lubridate)
library (reshape2)
library (gam)

source ("support.R")
```

```{r loadDataSets}
train.df <- read.csv ("data/train.csv")
test.df <- read.csv ("data/test.csv")

train.df <- formatData (train.df) %>% tbl_df()
test.df <- formatData (test.df) %>% tbl_df()
```

```{r plotAll}
long.df <- melt (train.df, measure.vars=c("count", "casual", "registered"), variable.name="type", value.name="total")

ggplot (long.df, aes (x=temp)) + geom_smooth (aes (y=total, color=type))
ggplot (long.df, aes (x=atemp)) + geom_smooth (aes (y=total, color=type))
ggplot (long.df, aes (x=windspeed)) + geom_smooth (aes (y=total, color=type))
ggplot (long.df, aes (x=humidity)) + geom_smooth (aes (y=total, color=type))
ggplot (long.df, aes (x=datetime)) + geom_smooth (aes (y=total, color=type))
ggplot (long.df, aes (x=hour)) + geom_smooth (aes(y=total, color=type))

ggplot (long.df, aes (x=wday, y=total, group=type, color=type)) + geom_point () +
        stat_smooth (method="gam", formula=y ~ ns(x, k=5))
 #+ geom_smooth (aes(color=type))
                                              
                                              
ggplot (long.df, aes (x=holiday, y=total)) + geom_boxplot() + facet_wrap(~type)
ggplot (long.df, aes (x=workingday, y=total)) + geom_boxplot() + facet_wrap(~type)
ggplot (long.df, aes (x=season, y=total, group=season)) + geom_boxplot() + facet_wrap (~type)
ggplot (long.df, aes (x=weather, y=total, group=weather)) + geom_boxplot() + facet_wrap (~type)
```

Conclusions:
1) Seems like ride count steadily increases with temperature until about 35C
2) Seems like atemp might be a better predictor compared to temp
3) Seems like windspeed may be a poor predictor
4) People don't like to rent when the humidity is high
5) In general, demand is dominated by regular users on working days, probably for their commute
6) Demand tends to be higher in the summer and fall
7) There tend to be huge outliers. The outliers are always high demand, rather than low demand


#rmarkdown::render ('exploratory.Rmd', output_file ='reports/explore.html')